\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[margin=1in]{geometry}

\begin{document}

\begin{center}
{\Large\bfseries Chapter 10, Problem 3}
\end{center}

\noindent\textbf{Problem.} Find the symmetry group, including only rotations, of a square, rectangular parallelepiped (i.e., square on the top and bottom and rectangular on the sides). Show that this group is isomorphic to the symmetry group of the square which includes both rotations and reflection. What are the classes? Show that there is an invariant subgroup. Find the factor group associated with this invariant subgroup.

\bigskip
\noindent\textbf{Solution.}

\noindent\textbf{Symmetry group of the square parallelepiped (rotations only):}

Consider a rectangular parallelepiped with a square cross-section (top and bottom are squares, sides are rectangles). The pure rotation symmetries are:

\begin{enumerate}
\item $e$: identity
\item $C_4$: rotation by $90^\circ$ about the vertical axis (through centers of square faces)
\item $C_4^2 = C_2$: rotation by $180^\circ$ about the vertical axis
\item $C_4^3$: rotation by $270^\circ$ about the vertical axis
\item $C_2'$: rotation by $180^\circ$ about a horizontal axis through midpoints of opposite square edges (say the $x$-axis)
\item $C_2''$: rotation by $180^\circ$ about the other horizontal axis through midpoints of opposite square edges (the $y$-axis)
\item $C_2^{(d)}$: rotation by $180^\circ$ about one diagonal axis of the square face
\item $C_2^{(d')}$: rotation by $180^\circ$ about the other diagonal axis of the square face
\end{enumerate}

This is a group of order 8, called $D_4$ (the dihedral group of order 8).

\medskip
\noindent\textbf{Symmetry group of the square (rotations and reflections):}

The symmetries of a planar square consist of:
\begin{enumerate}
\item $e$: identity
\item $r$: rotation by $90^\circ$
\item $r^2$: rotation by $180^\circ$
\item $r^3$: rotation by $270^\circ$
\item $\sigma_1, \sigma_2$: reflections through axes connecting midpoints of opposite sides
\item $\sigma_d, \sigma_{d'}$: reflections through the diagonals
\end{enumerate}

This is also the dihedral group $D_4$ of order 8.

\medskip
\noindent\textbf{Isomorphism:}

We construct the isomorphism $\phi$ by mapping:
\[
\phi(C_4) = r, \quad \phi(C_4^2) = r^2, \quad \phi(C_4^3) = r^3, \quad \phi(e) = e,
\]
\[
\phi(C_2') = \sigma_1, \quad \phi(C_2'') = \sigma_2, \quad \phi(C_2^{(d)}) = \sigma_d, \quad \phi(C_2^{(d')}) = \sigma_{d'}.
\]

Both groups have the same abstract structure: a cyclic subgroup of order 4 generated by the ``$90^\circ$'' element, and elements of order 2 outside this subgroup. The multiplication tables are identical under this correspondence, establishing the isomorphism.

\medskip
\noindent\textbf{Conjugacy Classes:}

The conjugacy classes of $D_4$ are:
\begin{align*}
&C_1 = \{e\}, \\
&C_2 = \{C_4, C_4^3\} \quad (\text{i.e., } \{r, r^3\}), \\
&C_3 = \{C_4^2\} \quad (\text{i.e., } \{r^2\}), \\
&C_4 = \{C_2', C_2''\} \quad (\text{i.e., } \{\sigma_1, \sigma_2\}), \\
&C_5 = \{C_2^{(d)}, C_2^{(d')}\} \quad (\text{i.e., } \{\sigma_d, \sigma_{d'}\}).
\end{align*}

There are 5 conjugacy classes.

\medskip
\noindent\textbf{Invariant Subgroup:}

The subgroup $N = \{e, C_4, C_4^2, C_4^3\} = \{e, r, r^2, r^3\}$ is a cyclic group of order 4. Since it is the union of conjugacy classes $C_1 \cup C_2 \cup C_3 = \{e, C_4, C_4^2, C_4^3\}$, it is an invariant (normal) subgroup.

Another invariant subgroup is $\{e, r^2, \sigma_1, \sigma_2\}$ and $\{e, r^2, \sigma_d, \sigma_{d'}\}$.

\medskip
\noindent\textbf{Factor Group:}

The factor group $D_4/N$ has two cosets:
\[
N = \{e, C_4, C_4^2, C_4^3\}, \quad C_2' N = \{C_2', C_2'', C_2^{(d)}, C_2^{(d')}\}.
\]

The factor group has order $8/4 = 2$ and is therefore isomorphic to $\mathbb{Z}_2$, with multiplication table:
\[
\begin{array}{c|cc}
 & N & C_2'N \\
\hline
N & N & C_2'N \\
C_2'N & C_2'N & N
\end{array}
\]

\end{document}
