\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[margin=1in]{geometry}

\begin{document}

\textbf{Chapter 6, Problem 10.} Find the Riemann surface on which $\sqrt{z^2 - 1} = \sqrt{z+1}\sqrt{z-1}$, $a$ real and positive, is a single-valued function, analytic except at $z = a$ and $z = -a$.

\bigskip
\textbf{Solution.}

Consider $f(z) = \sqrt{z^2 - a^2} = \sqrt{(z-a)(z+a)}$.

The branch points are at $z = a$ and $z = -a$.

\textbf{Branch cut selection.} We connect the two branch points with a branch cut along the real axis from $-a$ to $a$. Write:
\[
z + a = r_1 e^{i\theta_1}, \qquad z - a = r_2 e^{i\theta_2},
\]
where $r_1 = |z+a|$, $r_2 = |z-a|$.

With the branch cut along $[-a, a]$ on the real axis, we choose $\theta_1, \theta_2 \in (-\pi, \pi]$ (or equivalently $[0, 2\pi)$). The key property is that when we encircle both branch points together, both $\theta_1$ and $\theta_2$ increase by $2\pi$, so $(\theta_1 + \theta_2)/2$ increases by $2\pi$, and $e^{i(\theta_1+\theta_2)/2}$ returns to its original value. Hence $f$ is single-valued on a path encircling both branch points.

However, encircling only one branch point (say $z = a$) increases $\theta_2$ by $2\pi$ while $\theta_1$ remains essentially unchanged, giving $f \to -f$.

\textbf{Riemann surface.} We need a two-sheeted Riemann surface:
\begin{enumerate}
\item Take two copies of the complex plane, each cut from $-a$ to $a$ along the real axis.
\item On sheet 1: $f(z) = \sqrt{r_1 r_2}\, e^{i(\theta_1 + \theta_2)/2}$
\item On sheet 2: $f(z) = -\sqrt{r_1 r_2}\, e^{i(\theta_1 + \theta_2)/2}$
\item Join the upper lip of the cut on sheet 1 to the lower lip on sheet 2, and vice versa.
\end{enumerate}

On this surface, $f(z)$ is single-valued and analytic everywhere except at the branch points $z = \pm a$.

\textbf{The case $a = 1$:} $f(z) = \sqrt{z^2-1}$. The Riemann surface has:
\begin{itemize}
\item Two sheets joined along the cut $[-1, 1]$.
\item $f$ is analytic on both sheets except at $z = \pm 1$.
\item At infinity, $f(z) \sim z$ on one sheet and $f(z) \sim -z$ on the other.
\end{itemize}

\end{document}
