\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[margin=1in]{geometry}

\begin{document}

\textbf{Chapter 6, Problem 11.} Find the Riemann surface on which $\sqrt[n]{z^n - 1}$ is a single-valued function, analytic except at $z = e^{2\pi ik/n}$, $k = 0, 1, \ldots, n-1$.

\bigskip
\textbf{Solution.}

We have $f(z) = (z^n - 1)^{1/n}$. The branch points are the $n$-th roots of unity:
\[
z_k = e^{2\pi i k/n}, \qquad k = 0, 1, \ldots, n-1.
\]

Factor: $z^n - 1 = \prod_{k=0}^{n-1}(z - z_k)$.

So $f(z) = \left[\prod_{k=0}^{n-1}(z-z_k)\right]^{1/n}$.

\textbf{Analysis of branching.} Write $z - z_k = r_k e^{i\theta_k}$. Then:
\[
f(z) = \left(\prod r_k\right)^{1/n} \exp\!\left(\frac{i\sum_{k=0}^{n-1}\theta_k}{n}\right).
\]

Encircling a single branch point $z_j$: $\theta_j \to \theta_j + 2\pi$, all other $\theta_k$ unchanged. The total phase changes by $e^{2\pi i/n}$, so $f$ changes by a factor $e^{2\pi i/n} \neq 1$. Thus each $z_k$ is indeed a branch point.

Encircling all $n$ branch points: all $\theta_k \to \theta_k + 2\pi$, so the total phase changes by $e^{2\pi i \cdot n/n} = e^{2\pi i} = 1$. The function returns to its original value.

\textbf{Riemann surface construction.}
\begin{enumerate}
\item Take $n$ copies (sheets) of the complex plane.
\item Place branch cuts connecting the branch points. A convenient choice is to connect them in sequence around the unit circle, or to draw cuts from each $z_k$ to the origin (or to $\infty$). A particularly clean choice: connect $z_0$ to $z_1$, $z_2$ to $z_3$, etc.\ with straight cuts (for $n$ even), or draw $n$ cuts from each $z_k$ to the origin.

For the most symmetric arrangement with $n$ branch points, we can draw a single cut that passes through all of them---e.g., the unit circle itself serves as the branch cut, or radial cuts from each $z_k$ to infinity.

\item On each sheet $j$ ($j = 0, \ldots, n-1$), define:
\[
f_j(z) = \left(\prod r_k\right)^{1/n} \exp\!\left(\frac{i(\sum \theta_k + 2\pi j)}{n}\right).
\]

\item Join sheet $j$ to sheet $j+1 \pmod{n}$ across each branch cut.
\end{enumerate}

The result is an $n$-sheeted Riemann surface on which $f(z) = (z^n-1)^{1/n}$ is single-valued and analytic everywhere except at the branch points $z_k = e^{2\pi ik/n}$.

Note: at infinity, $f(z) \sim z$ (on the principal sheet), so infinity is a regular point of $f$ (not a branch point), consistent with the fact that encircling all branch points returns $f$ to its original value.

\end{document}
