\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[margin=1in]{geometry}

\begin{document}

\textbf{Chapter 6, Problem 16.} Explain why one of the following definite integrals is meaningful and the other meaningless. Evaluate the one that makes sense, writing the answer in the form $a + bi$:
\[
\text{a) } \int_i^{2i} \bar{z}\,dz, \qquad \text{b) } \int_0^i \sin 2z\,dz.
\]

\bigskip
\textbf{Solution.}

\textbf{(a)} $\int_i^{2i}\bar{z}\,dz$: The integrand $\bar{z}$ is \emph{not} an analytic function (it fails the Cauchy--Riemann equations). Therefore, the value of the integral depends on the path chosen from $i$ to $2i$. Without specifying a path, the integral is ambiguous.

However, if we take the straight-line path from $i$ to $2i$ (which is the natural interpretation), parametrize as $z = it$, $t \in [1,2]$, $dz = i\,dt$:
\[
\int_i^{2i}\bar{z}\,dz = \int_1^2 \overline{it}\cdot i\,dt = \int_1^2(-it)(i)\,dt = \int_1^2 t\,dt = \frac{t^2}{2}\Big|_1^2 = \frac{3}{2}.
\]

So on the straight-line path, the answer is $3/2$. But since $\bar{z}$ is not analytic, this integral is path-dependent and hence \emph{ambiguous without a specified path}.

\textbf{(b)} $\int_0^i \sin 2z\,dz$: The function $\sin 2z$ is an entire function (analytic everywhere). By Cauchy's theorem, the integral is path-independent. We can evaluate it directly:
\[
\int_0^i \sin 2z\,dz = \left[-\frac{\cos 2z}{2}\right]_0^i = -\frac{\cos 2i}{2} + \frac{\cos 0}{2}.
\]

Now $\cos(2i) = \cosh 2 = \frac{e^2 + e^{-2}}{2}$. Therefore:
\[
\int_0^i \sin 2z\,dz = \frac{1}{2} - \frac{\cosh 2}{2} = \frac{1 - \cosh 2}{2}.
\]

Numerically: $\cosh 2 = (e^2 + e^{-2})/2 \approx 3.7622$, so:
\[
\int_0^i \sin 2z\,dz = \frac{1 - 3.7622}{2} \approx -1.3811.
\]

This is a real number: $\boxed{\dfrac{1-\cosh 2}{2}}$.

The integral in (b) is the meaningful one (path-independent due to analyticity), while (a) is ambiguous without a specified path.

\end{document}
