\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[margin=1in]{geometry}

\begin{document}

\textbf{Chapter 6, Problem 22.} \textit{Conformal mapping.} Consider the analytic function $w = f(z)$ as transforming the complex $z$-plane into a complex $w$-plane. Suppose that two curves in the $z$-plane, $C_1$ and $C_2$, meet at an angle $\alpha$ at the point $z_0$ in the $z$-plane. Prove that the transformed curves in the $w$-plane, $C_1'$ and $C_2'$, meet at the same angle in the $w$-plane if $f'(z_0) \neq 0$. The requirement that $f'(z_0) \neq 0$ is essential for conformality at $z_0$. Consider the analytic function $w = z^2$, for which $w'(0) = 0$. The coordinate axes, which are separated by an angle of $90°$, are mapped into lines that are separated by $180°$. The angle between any two straight lines through the origin is doubled in the $w$-plane. Thus the mapping is not conformal at the origin. However, it is conformal everywhere else.

\bigskip
\textbf{Solution.}

Let two smooth curves $C_1$ and $C_2$ pass through the point $z_0$ with tangent directions $e^{i\alpha_1}$ and $e^{i\alpha_2}$, respectively, where $\alpha_1$ and $\alpha_2$ are the angles the tangent vectors make with the positive real axis. The angle between the curves at $z_0$ is $\alpha = \alpha_2 - \alpha_1$.

Parametrize the curves: $z_1(t)$ along $C_1$ with $z_1(0) = z_0$, and $z_2(t)$ along $C_2$ with $z_2(0) = z_0$. The tangent directions are $z_1'(0) = |z_1'(0)|e^{i\alpha_1}$ and $z_2'(0) = |z_2'(0)|e^{i\alpha_2}$.

Under the mapping $w = f(z)$, the image curves are $w_j(t) = f(z_j(t))$, $j = 1,2$.

By the chain rule:
\[
w_j'(0) = f'(z_0)\cdot z_j'(0).
\]

Write $f'(z_0) = |f'(z_0)|e^{i\phi}$ where $\phi = \arg f'(z_0)$. Then:
\[
w_j'(0) = |f'(z_0)|\cdot|z_j'(0)|\cdot e^{i(\phi + \alpha_j)}.
\]

The tangent angle of the image curve $C_j'$ in the $w$-plane is $\phi + \alpha_j$.

The angle between $C_1'$ and $C_2'$ in the $w$-plane:
\[
(\phi + \alpha_2) - (\phi + \alpha_1) = \alpha_2 - \alpha_1 = \alpha.
\]

The angle is preserved, and the rotation $\phi$ is the same for both curves. This proves the mapping is conformal at $z_0$, provided $f'(z_0) \neq 0$.

\textbf{The case $w = z^2$:} Here $f'(z) = 2z$, so $f'(0) = 0$. Near the origin, a curve making angle $\alpha$ with the real axis maps to a curve making angle $2\alpha$ (since $z = re^{i\theta} \mapsto w = r^2 e^{2i\theta}$). The positive real axis ($\theta = 0$) maps to $\theta_w = 0$, and the positive imaginary axis ($\theta = \pi/2$) maps to $\theta_w = \pi$. The $90°$ angle becomes $180°$. Angles are doubled, not preserved. Thus the mapping is not conformal at $z = 0$.

\end{document}
