\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[margin=1in]{geometry}

\begin{document}

\section*{Chapter 7, Problem 10}

\textbf{Problem:} Schr\"odinger's equation in one dimension is
\[
-\frac{\hbar^2}{2m}\frac{d^2\psi}{dx^2} + V(x)\psi(x) = E\psi(x).
\]
Using the methods discussed in Section 7.4, derive a Lippmann--Schwinger integral equation for $\psi(x)$,
\begin{itemize}
\item[a)] for the case $E < 0$,
\item[b)] for the case $E \ge 0$.
\end{itemize}
In case (b) you will find separate equations for $\psi_1(x)$ and $\psi_2(x)$ corresponding to different types of boundary conditions at infinity. Discuss the meaning of $\psi_1(x)$ and $\psi_2(x)$.

Solve Problem 10 by the method of Section 7.3. What is the appropriate statement of boundary conditions to specify the Green's function completely in this case?

\bigskip
\textbf{Solution:}

Rewrite Schr\"odinger's equation as:
\[
\frac{d^2\psi}{dx^2} + k^2\psi = \frac{2m}{\hbar^2}V(x)\psi(x),
\]
where $k^2 = 2mE/\hbar^2$.

The Green's function satisfies:
\[
\frac{d^2G}{dx^2} + k^2 G = \delta(x - x').
\]

\subsection*{(a) Case $E < 0$: $k^2 < 0$}

Let $\kappa^2 = -k^2 = 2m|E|/\hbar^2 > 0$, so $\kappa > 0$. The equation becomes:
\[
\frac{d^2G}{dx^2} - \kappa^2 G = \delta(x-x').
\]

The two independent solutions of the homogeneous equation are $e^{\kappa x}$ and $e^{-\kappa x}$. For the Green's function on $(-\infty,\infty)$, we require $G \to 0$ as $|x| \to \infty$:
\[
G(x,x') = \begin{cases} A\,e^{-\kappa x}, & x > x', \\ B\,e^{\kappa x}, & x < x'. \end{cases}
\]

Continuity at $x = x'$: $Ae^{-\kappa x'} = Be^{\kappa x'}$.

Jump in derivative: $-A\kappa e^{-\kappa x'} - B\kappa e^{\kappa x'} = 1$.

Solving: $A = -\frac{1}{2\kappa}e^{\kappa x'}$, $B = -\frac{1}{2\kappa}e^{-\kappa x'}$.

\[
\boxed{G(x,x') = -\frac{1}{2\kappa}e^{-\kappa|x-x'|}.}
\]

The Lippmann--Schwinger equation is:
\[
\psi(x) = \psi_0(x) + \int_{-\infty}^{\infty} G(x,x')\frac{2m}{\hbar^2}V(x')\psi(x')\,dx',
\]
where $\psi_0$ is a solution of the free equation. For $E < 0$, the only normalizable free solution is $\psi_0 = 0$, so:
\[
\psi(x) = -\frac{m}{\hbar^2\kappa}\int_{-\infty}^{\infty} e^{-\kappa|x-x'|}V(x')\psi(x')\,dx'.
\]

\subsection*{(b) Case $E \ge 0$: $k^2 > 0$}

Now $k > 0$ is real. The homogeneous solutions are $e^{\pm ikx}$. There is no unique Green's function without specifying boundary conditions. Two choices are:

\textbf{Outgoing-wave (retarded) Green's function:}
\[
G_+(x,x') = -\frac{1}{2ik}e^{ik|x-x'|}.
\]
This gives outgoing waves at $|x| \to \infty$.

\textbf{Incoming-wave (advanced) Green's function:}
\[
G_-(x,x') = \frac{1}{2ik}e^{-ik|x-x'|}.
\]

The corresponding Lippmann--Schwinger equations are:
\[
\psi_{\pm}(x) = e^{ikx} + \frac{2m}{\hbar^2}\int_{-\infty}^{\infty}G_{\pm}(x,x')V(x')\psi_{\pm}(x')\,dx'.
\]

Here $\psi_0 = e^{ikx}$ represents an incoming plane wave from the left.

\textbf{Physical interpretation:} $\psi_+$ describes scattering with outgoing scattered waves (the physical scattering solution), while $\psi_-$ describes scattering with incoming scattered waves (time-reversed scattering).

\subsection*{Method of Section 7.3}

By the eigenfunction expansion method, the Green's function is:
\[
G(x,x') = \int_{-\infty}^{\infty}\frac{e^{ip(x-x')/\hbar}}{k^2 - p^2/\hbar^2}\,\frac{dp}{2\pi\hbar}.
\]

The poles at $p = \pm\hbar k$ require a prescription for the contour. For the retarded Green's function, we use $k \to k + i\epsilon$, pushing the poles off the real axis. This uniquely specifies the boundary conditions (outgoing waves) and gives $G_+$ after contour integration. For the advanced Green's function, $k \to k - i\epsilon$. For $E < 0$, the poles are on the imaginary axis and no prescription is needed. $\blacksquare$

\end{document}
