\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[margin=1in]{geometry}

\begin{document}

\section*{Chapter 7, Problem 25}

\textbf{Problem:}
\begin{itemize}
\item[a)] Show that the function
\[
G_t(\mathbf{r},t;\mathbf{r}',t') = -G(\mathbf{r},t;\mathbf{r}',t')\,\theta(t'-t)
\]
is a Green's function for Eq.\ (7.131), where $G_t$ satisfies Eqs.\ (7.134) and (7.135).

This means that we can write a time-dependent Schr\"odinger integral equation as:
\[
\psi(\mathbf{r},t) = \psi_0(\mathbf{r},t) - \frac{i}{\hbar}\int_{-\infty}^{t}dt'\int G_t(\mathbf{r},t;\mathbf{r}',t')V(\mathbf{r}',t')\psi(\mathbf{r}',t')\,d^3r',
\]
in the notation of Eq.\ (7.137). Show that $U$ is a static potential which is ``turned off'' at remote future times, that is, $V \to V_0 e^{-\epsilon|t|}$, then we obtain in the limit $t \to 0$ the Lippmann--Schwinger equation with an incoming spherical wave boundary condition solution.
\item[b)] Note that Eq.\ (7.129a) does not make sense if $t' > t$ (assuming, as will usually be the case, that $\lambda_n \to +\infty$ as $n \to \infty$ and $t$ and $t'$ are real). This is also reflected in the discussion following Eq.\ (7.128), where we are tacitly assuming that $t = 0$ from ``the positive side of zero'' (see Section 5.3). Thus the Green's function
\[
G(\mathbf{r},t;\mathbf{r}',t') = \sum_n \phi_n(\mathbf{r})\phi_n^*(\mathbf{r}')\,e^{-i\lambda_n(t-t')}
\]
obtained in the previous problem, although it formally seems to satisfy
\[
\Bigl(H + \frac{i\partial}{\partial t}\Bigr)G = \delta^3(\mathbf{r}-\mathbf{r}')\delta(t-t'),
\]
is not an acceptable Green's function. Show that it does give $G_t(\mathbf{r},t;\mathbf{r}',t')$ for the case of the diffusion (or heat) equation by using the Fourier transform method, one obtains only
\[
G(\mathbf{r},t;\mathbf{r}',t') = \theta(t-t')\,G_{\mathrm{ret}}(\mathbf{r},t;\mathbf{r}',t'),
\]
where $G_t$ is given by Eq.\ (7.139).
\end{itemize}

\bigskip
\textbf{Solution:}

\subsection*{Part (a)}

The time-dependent Schr\"odinger equation is:
\[
i\hbar\frac{\partial\psi}{\partial t} = (H_0 + V)\psi.
\]

The retarded Green's function satisfies:
\[
\Bigl(i\hbar\frac{\partial}{\partial t} - H_0\Bigr)G_{\mathrm{ret}}(\mathbf{r},t;\mathbf{r}',t') = \delta^3(\mathbf{r}-\mathbf{r}')\delta(t-t'),
\]
with the causal boundary condition $G_{\mathrm{ret}} = 0$ for $t < t'$.

The advanced Green's function $G_t = -G\,\theta(t'-t)$ satisfies $G_t = 0$ for $t > t'$. To verify it satisfies the adjoint equation:
\[
\Bigl(i\hbar\frac{\partial}{\partial t} - H_0\Bigr)G_t = \Bigl(i\hbar\frac{\partial}{\partial t} - H_0\Bigr)\bigl[-G\,\theta(t'-t)\bigr].
\]

Using $\partial\theta(t'-t)/\partial t = -\delta(t-t')$ and the fact that $G$ satisfies the Schr\"odinger equation for $t \neq t'$:
\[
= -\bigl[\delta^3(\mathbf{r}-\mathbf{r}')\delta(t-t')\bigr]\theta(t'-t) + i\hbar\,G\,\delta(t-t') - (H_0 G)(-\theta(t'-t)).
\]

At $t = t'$, $G(\mathbf{r},t;\mathbf{r}',t) = \delta^3(\mathbf{r}-\mathbf{r}')$ (the initial condition). So:
\[
= -\delta^3(\mathbf{r}-\mathbf{r}')\delta(t-t').
\]

For the Lippmann--Schwinger connection: with $V \to V_0 e^{-\epsilon|t|}$ and using $G_t$ (the advanced Green's function), the integral equation in the $t\to\infty$ limit picks up the incoming-wave boundary condition, which is the time-reversed version of the outgoing-wave condition obtained with the retarded Green's function.

\subsection*{Part (b)}

The eigenfunction expansion $G = \sum_n \phi_n(\mathbf{r})\phi_n^*(\mathbf{r}')e^{-i\lambda_n(t-t')}$ converges only for $t > t'$ (when eigenvalues correspond to $\lambda_n > 0$, giving exponential damping in the diffusion case). For $t < t'$, the series diverges.

Using the Fourier transform in time, the retarded Green's function for the diffusion equation $(\partial_t - k\nabla^2)G = \delta$ is obtained by closing the contour in the lower half-plane (for $t > t'$):
\[
G_{\mathrm{ret}}(\mathbf{r},t;\mathbf{r}',t') = \theta(t-t')\frac{1}{[4\pi k(t-t')]^{d/2}}\exp\!\Bigl(-\frac{|\mathbf{r}-\mathbf{r}'|^2}{4k(t-t')}\Bigr),
\]
which automatically includes the $\theta(t-t')$ factor from causality. The Fourier method naturally produces only the retarded Green's function because the diffusion equation is first-order in time and dissipative. $\blacksquare$

\end{document}
