\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[margin=1in]{geometry}

\begin{document}

\section*{Chapter 3, Problem 22}

\textbf{Problem.} Consider an $n \times n$ matrix $Q$ of a quantum oscillator. Show that all eigenvalues of the quantum oscillator have the form $E_n = \hbar\omega(n + \tfrac{1}{2})$.

\textit{[Hint: Assume that there is an eigenvector $v$ and associated eigenvalue $\lambda$ such that $Hv = \lambda v$ and using $(a, a^\dagger) = \hbar\omega$, create $E_n = \hbar\omega n$, etc., and arrive at eigenvalues.]}

\bigskip
\textbf{Solution.}

The quantum harmonic oscillator Hamiltonian is
\[
H = \frac{p^2}{2m} + \frac{1}{2}m\omega^2 x^2.
\]
Introducing the ladder operators:
\[
a = \sqrt{\frac{m\omega}{2\hbar}}\left(x + \frac{ip}{m\omega}\right), \qquad a^\dagger = \sqrt{\frac{m\omega}{2\hbar}}\left(x - \frac{ip}{m\omega}\right),
\]
with the commutation relation $[a, a^\dagger] = 1$, the Hamiltonian becomes:
\[
H = \hbar\omega\left(a^\dagger a + \tfrac{1}{2}\right).
\]

\textbf{Step 1: Show eigenvalues are bounded below.}

For any state $|\psi\rangle$:
\[
\langle \psi | H | \psi \rangle = \hbar\omega\left(\langle \psi | a^\dagger a | \psi \rangle + \tfrac{1}{2}\right) = \hbar\omega\left(\|a|\psi\rangle\|^2 + \tfrac{1}{2}\right) \ge \tfrac{1}{2}\hbar\omega.
\]
So all eigenvalues satisfy $E \ge \tfrac{1}{2}\hbar\omega$.

\textbf{Step 2: Action of ladder operators on eigenstates.}

Suppose $H|v\rangle = E|v\rangle$. Using $[H, a^\dagger] = \hbar\omega\, a^\dagger$ and $[H, a] = -\hbar\omega\, a$:

\[
H(a^\dagger|v\rangle) = (a^\dagger H + \hbar\omega\, a^\dagger)|v\rangle = (E + \hbar\omega)(a^\dagger|v\rangle).
\]
So $a^\dagger|v\rangle$ is an eigenvector with eigenvalue $E + \hbar\omega$ (the raising operator).

\[
H(a|v\rangle) = (aH - \hbar\omega\, a)|v\rangle = (E - \hbar\omega)(a|v\rangle).
\]
So $a|v\rangle$ is an eigenvector with eigenvalue $E - \hbar\omega$ (the lowering operator).

\textbf{Step 3: Ground state.}

Applying $a$ repeatedly lowers the eigenvalue by $\hbar\omega$ each time. Since $E \ge \tfrac{1}{2}\hbar\omega$, this process must terminate. It terminates when $a|0\rangle = 0$ for some ground state $|0\rangle$. Then:
\[
H|0\rangle = \hbar\omega\left(a^\dagger a + \tfrac{1}{2}\right)|0\rangle = \hbar\omega\left(0 + \tfrac{1}{2}\right)|0\rangle = \tfrac{1}{2}\hbar\omega\,|0\rangle.
\]
So $E_0 = \tfrac{1}{2}\hbar\omega$.

\textbf{Step 4: All eigenvalues.}

Applying $(a^\dagger)^n$ to the ground state produces states with eigenvalue:
\[
E_n = E_0 + n\hbar\omega = \hbar\omega\left(n + \tfrac{1}{2}\right), \qquad n = 0, 1, 2, \ldots
\]

No other eigenvalues exist because any eigenvector can be lowered to the ground state by repeated application of $a$, and the lowering steps are quantized in units of $\hbar\omega$. \hfill $\blacksquare$

\end{document}
